---
title: "DDS Case Study 2"
author: "David Julovich"
date: "8/15/2019"
output: html_document
---

```{r, include=FALSE}

library(class)
library(gmodels)
library(caret)
library(magrittr)
library(ROSE)
library(tidyverse) #general data wrangling tools
library(skimr) #summary stats
library(corrplot)
library(MASS)#studentized residuals
library(lindia)
library(ggplot2)
library(car) # Regression tools
library(rlang)
library(olsrr)
library(forcats) 
library(roperators)
library(gridExtra)
library(mefa4)
library(stats)
library(caret)
library(tidyverse) #general data wrangling tools
library(skimr) #summary stats
library(corrplot)
library(MASS)#studentized residuals
library(lindia)
library(ggplot2)
library(car) # Regression tools
library(rlang)
library(olsrr)
library(forcats) 
library(roperators)
library(gridExtra)
library(mefa4)
library(stats)
```

```{r  echo=TRUE}

cs2<- read.csv("c:/Users/daj0079/Desktop/SMU/CaseStudy2-data.csv")
no_at<-read.csv("c:/Users/daj0079/Desktop/SMU/CaseStudy2Validation No Attrition.csv")

#Normalize for KNN
normalize <-function(x){
   return((x-min(x))/(max(x)-min(x)))}

#Set up the data for the model 
#change data to factors and remove useless variable

OT<-as.numeric(cs2$OverTime)
OTO<-as.numeric(no_at$OverTime)
no_at<-cbind(no_at,OTO)
cs2<-cbind(cs2, OT)
cs2.sub<-cs2[c( 'Attrition', 'JobLevel', 'OT')]
table(cs2$Attrition)
#Run Normalize and Build tain and test groups
cs2_sub.norm<-as.data.frame(lapply(cs2.sub[-1], normalize))
set.seed(123)
trainobs=sample(seq(1,dim(cs2_sub.norm)[1]),round(.75*dim(cs2_sub.norm)[1]),replace=FALSE)
cs_Train =cs2_sub.norm[trainobs,]
cs_Test=cs2_sub.norm[-trainobs,]
cs_train_target <-cs2[trainobs,3]
cs_test_target<-cs2[-trainobs,3]

#1 Model 1 knn K=3
model_1<-knn(train=cs_Train, test=cs_Test,cl= cs_train_target, k=3)
#table and confusion matrix
CrossTable(x=cs_test_target,y=model_1, prop.chisq=FALSE)
confusionMatrix(model_1,cs_test_target, dnn = c("Prediction", "Reference"),  mode = "sens_spec")




#2 Over sampling model for attrition = YES k=12
data_balanced_over <- ovun.sample(Attrition ~ ., data = cs2, method = "over",N = 1200)$data
table(data_balanced_over$Attrition)

#Set up the data for the model
data_over.sub<-data_balanced_over[c( 'Attrition', 'JobLevel', 'OT','Age', 'Education', 'JobSatisfaction',  'DistanceFromHome', 'JobInvolvement', 'YearsAtCompany','NumCompaniesWorked')]

#Run Normalize and Build tain and test groups
data_over.norm<-as.data.frame(lapply(data_over.sub[-1], normalize))
set.seed(899)
data_over_trainobs=sample(seq(1,dim(data_over.norm)[1]),round(.75*dim(data_over.norm)[1]),replace=FALSE)
over_cs_Train =data_over.norm[data_over_trainobs,]
over_cs_Test=data_over.norm[-data_over_trainobs,]
over_cs_train_target <-data_balanced_over[data_over_trainobs,3]
over_cs_test_target<-data_balanced_over[-data_over_trainobs,3]

#knn k=12
over_sam<-knn(train=over_cs_Train, test=over_cs_Test,cl= over_cs_train_target, k=12)
#table and confusion matrix
CrossTable(x=over_cs_test_target,y=over_sam, prop.chisq=FALSE)
confusionMatrix(over_sam,over_cs_test_target, dnn = c("Prediction", "Reference"),  mode = "sens_spec")




#3 Attrition Predicition Model k=12
no_at.sub<-no_at[c( 'JobLevel', 'OTO','Age', 'Education', 'JobSatisfaction',  'DistanceFromHome', 'JobInvolvement', 'YearsAtCompany','NumCompaniesWorked')]

#Run Normalize and Build tain and test groups
no_at.norm<-as.data.frame(lapply(no_at.sub, normalize))

#knn k=12
no_at_pred<-knn(train=over_cs_Train, test=no_at.norm,cl= over_cs_train_target, k=12)

#table of predictions
no_at_pred
CrossTable(x=over_cs_test_target,y=no_at_pred, prop.chisq=FALSE)



rcs2<- read.csv("c:/Users/daj0079/Desktop/SMU/CaseStudy2-data.csv")
no_sal<-read.csv("c:/Users/daj0079/Desktop/SMU/CaseStudy2Validation No Salary.csv")
#yes
ggplot(rcs2, aes(x = Attrition, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#yes
ggplot(rcs2, aes(x = BusinessTravel, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#yes
ggplot(rcs2, aes(x = Department, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#YES
ggplot(rcs2, aes(x =EducationField, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = Gender, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#YES
ggplot(rcs2, aes(x = JobRole, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#yes
ggplot(rcs2, aes(x = MaritalStatus, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = Over18, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
ggplot(rcs2, aes(x = OverTime, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
##YES
ggplot(rcs2, aes(x = Age, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
##NO
ggplot(rcs2, aes(x = DailyRate, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
##NO
ggplot(rcs2, aes(x = DistanceFromHome, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#Yes
ggplot(rcs2, aes(x = Education, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = EmployeeCount, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = EnvironmentSatisfaction, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = HourlyRate, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = ID, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = JobInvolvement, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
###YES#### make factor
ggplot(rcs2, aes(x = JobLevel, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)+ggtitle("Regression")+theme(plot.title = element_text(hjust = 0.5))
#NO
ggplot(rcs2, aes(x = JobSatisfaction, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = MonthlyRate, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = NumCompaniesWorked, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = PercentSalaryHike, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = PerformanceRating, y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = RelationshipSatisfaction,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = TrainingTimesLastYear,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
#NO
ggplot(rcs2, aes(x = WorkLifeBalance,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
####YES FACTOR###
ggplot(rcs2, aes(x = YearsAtCompany,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
####YES FACTOR###
ggplot(rcs2, aes(x = YearsInCurrentRole,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
####YES FACTOR###
ggplot(rcs2, aes(x = YearsSinceLastPromotion,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)
###YES FACTOR###
ggplot(rcs2, aes(x = YearsWithCurrManager,y = MonthlyIncome)) +
   geom_point() + geom_smooth(method=lm)


rcs2 %<>% mutate(logdistance = log(DistanceFromHome)) %>% mutate(logmonthlyincome = log(MonthlyIncome))%>%mutate(logyearsatcompany=log(YearsAtCompany))
no_sal$Attrition<-ifelse(no_sal$Attrition=="Yes",1, ifelse(no_sal$Attrition =="No",0,0))
rcs2$Attrition<-ifelse(rcs2$Attrition=="Yes",1, ifelse(rcs2$Attrition =="No",0,0))


rcs2$JobLevel<-as.factor(rcs2$JobLevel)
no_sal$JobLevel <- as.factor(no_sal$JobLevel)

rcs2$Gender<-ifelse(rcs2$Gender=="Male",1, ifelse(rcs2$Gender =="Female",2,0))
no_sal$OverTime<-ifelse(no_sal$OverTime=="Yes",1, ifelse(no_sal$OverTime =="No",2,0))


imp_drop<-rcs2[c("Age", "Education", "JobRole", 'JobLevel',  "TotalWorkingYears","MonthlyIncome")]
  



imp_drop<-rcs2[c("Age", "Education", "JobRole", 'JobLevel', "MaritalStatus",  "TotalWorkingYears","MonthlyIncome")]

no_sal_drop<-no_sal[c("Age", "Education", "JobRole", 'JobLevel', "MaritalStatus",  "TotalWorkingYears")]


#imp_testing=rcs2[,!(names(rcs2) %in% imp_drop)]


imp.model <- lm(MonthlyIncome ~., data = imp_drop)
summary(imp.model)
vif(imp.model)
plot(imp.model)

imp_pred<-predict(imp.model, imp_drop)


no_sal_pred<-predict(imp.model,no_sal_drop)
no_sal_pred


RMSE = function(m, o){
   sqrt(mean((m - o)^2))
}

score<-RMSE(imp_drop$MonthlyIncome,imp_pred)
score

